---
title: "Death Metal's development"
author: "Quang DO"
date: "10 février 2018"
output: 
  html_document:
    number_sections: true
    css: custom.css
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading data {.tabset}

```{r loading libraries, message = FALSE, warning = FALSE}
library(data.table)
library(plotly)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)

setwd("C:/documents/xq.do/Desktop/Kaggle/Death_Metal/")

albums = fread("./data/albums.csv", stringsAsFactors = T)
bands = fread("./data/bands.csv", stringsAsFactors = T)
reviews = fread("./data/reviews.csv", stringsAsFactors = T)

#-----      formatting     -----#
bands[, formed_in := as.numeric(as.character(formed_in))]
```

# Some simple statistics {.tabset}

## The first look

Death metal is a very big genre of metal, can can be considered as the most violent genre together with black metal. The first thing to do here is to look at how many death metal bands / albums there are in the world?

```{r first look}
print(paste0("There are ", dim(bands)[1], " death metal bands in the world"))
print(paste0("There are ", dim(albums)[1], " death metal albums in the world"))
```


## World's first metal band?
```{r first band}
first_band = bands[ , .SD[which.min(formed_in)]]
print(paste0("The first deth metal band is: ", first_band$name))
print(paste0("They were found in ", first_band$formed_in, " in ", first_band$country))
```

What? Wait, the first death metal band is not Possessed? Ah the problem here is that "Satan's Host" was found before Possessed and changed their genre afterwards.


## First album?
```{r first album}
first_album = albums[, .SD[which.min(year)]]
print(paste0("World's first death metal album is: ", first_album$title, " released in ", first_album$year))
```


# Cumulative number of bands and albums? {.tabset}

```{r cum nb of band}
bands[, nb.of.band := .N, by = formed_in]
band.count = unique(bands[, c("nb.of.band", "formed_in")])
band.count = band.count[order(formed_in)]
band.count = band.count[!is.na(formed_in), ]
band.count[, nb.of.band.cum := cumsum(nb.of.band)]

albums[, nb.of.album := .N, by = year]
album.count = unique(albums[, c("nb.of.album", "year")])
album.count = album.count[order(year)]
album.count = album.count[!is.na(year), ]
album.count[, nb.of.album.cum := cumsum(nb.of.album)]

band.album = merge(album.count, band.count, by.x = "year", by.y = "formed_in")

p1 = plot_ly(band.album, x = ~year, y = ~nb.of.band.cum, name = 'number of bands', type = "scatter", mode = 'lines+markers') %>%
  add_trace(y = ~nb.of.album.cum, name = 'number of albums', mode = 'lines+markers') %>%
  layout(yaxis = list(title = FALSE), legend = list(x = 0.1, y = 0.9))
p1
```

The fact that there are more bands than albums suggests that there are a lot of cover bands (bands withou any released album)


#Cover bands development?

So let's take a look at the development of the cover bands

```{r cover band}
#Number of bands without albums per year
cover.band = merge(bands, albums, by.x = "id", by.y = "band", all.x = T)
cover.band = cover.band[!is.na(year)][, nb.of.cover.band := .N, by = formed_in]
cover.band = unique(cover.band[order(formed_in), c("nb.of.cover.band", "formed_in")])

p3 = plot_ly(cover.band, x = ~formed_in, y = ~nb.of.cover.band, type = "scatter", mode = 'lines+markers') %>%
  layout(yaxis = list(title = FALSE), legend = list(x = 0.1, y = 0.9),
         title = 'number of new cover bands')
p3
```

# number of new bands? {.tabset}

```{r nb of new bands}
p2 = plot_ly(band.album, x = ~year, y = ~nb.of.band, name = 'number of new bands', type = "scatter", mode = 'lines+markers') %>%
  add_trace(y = ~nb.of.album, name = 'number of new albums', mode = 'lines+markers') %>%
  layout(yaxis = list(title = FALSE), legend = list(x = 0.1, y = 0.9))
p2
```

We can see that 2005 is the prime year of forming a death metal band.


# Death metal subgenre

what are the main subgenre of death metal? what is the distribution of its subgenres?

```{r dm subgenres}
#Distribution of main genre?
#Distribution of main genre?
bands[, genre := as.character(genre)]
bands[, genre1 := sapply(genre, function(x) strsplit(x, '|', fixed = T)[[1]][1])]
bands[, genre2 := sapply(genre, function(x) strsplit(x, '|', fixed = T)[[1]][2])]
bands[, genre3 := sapply(genre, function(x) strsplit(x, '|', fixed = T)[[1]][3])]


genre.count = bands[, list('count' = sum(genre != '')), by = genre1]
genre.count = genre.count[order(-count)]
genre.count[, genre1 := ifelse(count <= genre.count$count[11], "others", genre1)]


p5 = plot_ly(labels = ~genre.count$genre1, values = ~genre.count$count) %>%
     add_pie(hole = 0.6) %>%
     layout(title = "Distribution of death metal subgenre",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
p5
```



# Fastest developped genres

What sub-genres of Death Metal develop faster more than the other?

```{r subgenre development}
#Which subgenre develops faster?
main.genre = unique(genre.count$genre1)
main.genre = main.genre[1:8]


main.genre.bands = bands[genre1 %in% main.genre, c("id", "genre1")]
main.genre.albums = merge(albums, main.genre.bands, 
                          by.x = "band", by.y = "id",
                          all.x = T)
main.genre.albums = main.genre.albums[!is.na(genre1),]

main.genre.dev = main.genre.albums[, list("count.main.genre.album" = .N),
                                    by = c("year", "genre1")]

main.genre.dev = main.genre.dev[order(genre1, year)]
main.genre.dev = main.genre.dev[year <= 2015, ]

p6 = plot_ly(x     = ~main.genre.dev$year, 
             y     = ~main.genre.dev$count.main.genre.album, 
             split = ~main.genre.dev$genre1,
             type  = "scatter",
             color = ~main.genre.dev$genre1, 
             mode  = "lines+markers") %>%
     layout(yaxis = list(title = FALSE), legend = list(x = 0.1, y = 1),
            title = "number of new albums by genre")

p6
```




# Most productive bands

The productivity here is defined by the number of album a band releases

```{r productive bands}
#Which bands are most productive in terms of number of albums?
album.count.by.bands = albums[, list("album.count" = .N),
                              by = band]

album.count.by.bands = merge(album.count.by.bands, bands[, c("id", "name")],
                             by.x = "band", by.y = "id")

album.count.by.bands = album.count.by.bands[order(-album.count)]
album.count.by.bands = album.count.by.bands[1:15, ]
album.count.by.bands[, name := droplevels(name)]

p7 = plot_ly(x = album.count.by.bands$album.count, 
             y = reorder(album.count.by.bands$name, album.count.by.bands$album.count), 
             type = 'bar', 
             orientation = 'h') %>%
    layout(margin = list(l = 200))
  
p7
```


Cirle is the most productive band with 31 albums in total, followed by Morcegos and Transmetal with 23 albums each. I only know Darkthrone, Therion & Napalm Death in this list.

# Most popular albums

The popularity here is defined by the number of reviews an album receives on metal-archives.com

```{r popular album}
#which albums are most popular? (by number of reviews)
albums.band = merge(albums, bands[, c("name", "id")], by.x = "band", by.y = "id", all.x = T)
popu.albums = merge(albums.band, reviews[, c("album", "score")],
                    by.x = "id", by.y = "album",
                    all.x = T)
popu.albums[, name := paste0(name, " - '", title, "'")]
popu.albums = popu.albums[, list("nb.reviews" = .N), by = c("name", "title")][order(-nb.reviews), ]
popu.albums = popu.albums[1:20, ]

p8 = plot_ly(x = popu.albums$nb.reviews, 
             y = reorder(popu.albums$name, popu.albums$nb.reviews), 
             type = 'bar', 
             orientation = 'h') %>%
  layout(margin = list(l = 500))
p8
```


# Best albums

```{r best albums}
#best album?
best.albums = merge(albums.band, reviews[, c("album", "score")],
                    by.x = "id", by.y = "album",
                    all.x = T)
best.albums = best.albums[, name := paste0(name, " - '", title, "'")][!is.na(score), ]
best.albums = best.albums[, list("note" = round(mean(score, na.rm = T), 2),
                                 "nb.review" = .N), 
                          by = c("name")]
best.albums = best.albums[nb.review > 10, ][order(-note),][1:20, ]
p8 = plot_ly(x = best.albums$note, 
             y = reorder(best.albums$name, best.albums$note), 
             text = best.albums$note,
             textposition = 'auto',
             type = 'bar', 
             orientation = 'h') %>%
  layout(margin = list(l = 300))
p8
```


# Most metal countries

we can see the number of death metal bands in a country

```{r band country}
country.band.count = bands[, list("bands.count" = .N),
                           by = country][1:15, ]
p10 = plot_ly(x = country.band.count$bands.count, 
             y = reorder(country.band.count$country, country.band.count$bands.count), 
             type = 'bar', 
             orientation = 'h') %>%
  layout(margin = list(l = 300))
p10
```

# Lyrics {.tabset}

## helper funtion

We need to first define a helper function to use after.

```{r word cloud}
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
plot.me.the.cloud = function(text){
  docs = Corpus(text)
  docs <- tm_map(docs, toSpace, "/")
  docs <- tm_map(docs, toSpace, "@")
  docs <- tm_map(docs, toSpace, "\\|")
  
  # Convert the text to lower case
  docs <- tm_map(docs, content_transformer(tolower))
  # Remove numbers
  docs <- tm_map(docs, removeNumbers)
  # Remove english common stopwords
  docs <- tm_map(docs, removeWords, stopwords("english"))
  # Remove your own stop word
  # specify your stopwords as a character vector
  docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 
  # Remove punctuations
  docs <- tm_map(docs, removePunctuation)
  # Eliminate extra white spaces
  docs <- tm_map(docs, stripWhitespace)
  
  
  dtm <- TermDocumentMatrix(docs)
  m <- as.matrix(dtm)
  v <- sort(rowSums(m),decreasing=TRUE)
  d <- data.frame(word = names(v),freq=v)
  head(d, 10)
  
  set.seed(1234)
  wordcloud(words = d$word, freq = d$freq, min.freq = 1,
            max.words=50, random.order=FALSE, rot.per=0.35, 
            colors=brewer.pal(8, "Dark2"))
}
```

## overall

```{r}
text = VectorSource(bands[theme != "N/A", ]$theme)
plot.me.the.cloud(text)
```

## Melodic Death Metal

```{r}
text = VectorSource(bands[genre1 == "Melodic Death Metal" & theme != "N/A", ]$theme)
plot.me.the.cloud(text)
```

## Brutal Death Metal

```{r}
text = VectorSource(bands[genre1 == "Brutal Death Metal" & theme != "N/A", ]$theme)
plot.me.the.cloud(text)
```


## Black/Death Metal

```{r}
text = VectorSource(bands[genre1 == "Black/Death Metal" & theme != "N/A", ]$theme)
plot.me.the.cloud(text)
```


